---
title: "stringsAsFactors = HELLNO"
author: "Vibudh Agrawal"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stringsAsFactors = HELLNO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Factors are a very useful type of variable in R, but they can also drive you nuts. Especially the "stealth factor" that you think of as character.

Can we soften some of their sharp edges?

Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work.  Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```

Detecting whether a given factor should be character fact_detect():

Sometime factors are the same as characters as there are no redundant charaters so not required and might be difficult to check if the input is huge:

```{r}
a <- factor(c("this","is","test","case"))
print(a)
```

fact_detect() retuen true if factors are not (i.e. factors are characters) required and false if factors are required 

```{r}
fact_detect(a)
```


Converting the order of factors to given by user fact_reord():

Levels are always in the ascending order:
```{r}
a <- c("This","is","test","examp")
factor(a)
```

If we want to plot the order of plotting is based on levels and not data so we might want to reorder our levels:
```{r}
fact_reord(a,c(4,3,2,1))
```

Changing the order of levels to that of original character vector fact_orig():

Levels are always in the ascending order:
```{r}
a <- c("This","is","test","examp")
factor(a)
```

Sometimes we want the levels to be exactly that of our given input character vector:
```{r}
fact_orig(a)
```





